---
title: "Federal Employment Network"
author: "Blain Morin"
date: "4/7/2021"
runtime: shiny
output: 
    ioslides_presentation: 
      widescreen: yes
      transition: slower
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cahce = TRUE)

library(shiny)
library(readr)
library(tidyverse)
library(networkD3)
library(plotly)
library(readr)
library(statnet)
library(ndtv)
library(latticeExtra)

ref = read.csv("fed_agency_capacity_autonomy.csv")
fed = read.csv("r73_13_recoded_flows.csv")


fed = fed %>%
    filter(AGYSUB_next != "EXIT") %>%
    filter(AGYSUB_next != "UKNOWN")

```


# Introduction

## Background

* In the disaster literature, networks of actors form to address complex problem
* Prior studies look at the communication networks that form during these events
* *Capacity*: The resources and professionalization required to accomplish tasks and goals. 

## Motivation

```{r}

test = fed %>%
  filter(AGYSUB_next != "EXIT")

test = test %>%
  filter(AGYSUB == "EP00" | AGYSUB_next == "EP00")

test2 = test %>%
  group_by(yr) %>%
  summarize(Sent = sum(AGYSUB == "EP00"), Received = sum(AGYSUB_next == "EP00"))

test2 = test2 %>%
  gather(key = Measure, value = Edge, 2:3)


p = test2 %>%
  ggplot(aes(x = yr, y = Edge)) +
  geom_line(aes(group = Measure, color = Measure), size = 1.5) +
  ylab("Number of Employees") +
  xlab("Year") +
  ggtitle("Environmental Protection Agency: Employees Sent and Received") +
  theme_bw() +
  scale_color_brewer(type = "qual", palette = 2) +
  scale_x_continuous(breaks = c(1973, seq(1976, 2010, by = 4,), 2012), minor_breaks = 1973:2012) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  
ggplotly(p) %>%
    config(displayModeBar = FALSE)

```

# Data

## Source

* Buzzfeed FOAI request

## Network Visualization 

```{r}

fluidRow(
        column(3,
            numericInput("year",
                        "Year:",
                        min = 1974,
                        max = 2012,
                        value = 1974)),
        column(3,    
            sliderInput("nmin",
                         "Min Agency Size:",
                        min = 1, max = max(fed$n),
                        value = 10000)),
        column(2,    
               numericInput("send",
                           "Min. Sent:",
                           min = 1, max = 2000,
                           value = 10, step = 5)),
            
        column(2,    
            checkboxInput("EPA", "EPA", value = TRUE))
    )

```

```{r}

renderForceNetwork({
        # generate bins based on input$bins from ui.R
    
        if (input$EPA == TRUE){
            df = fed %>%
            filter(yr == input$year) %>%
            filter(n >= input$nmin) %>%
            filter(n_next >= input$send[1], n_next <= 2000) %>%
            filter(AGYSUB == "EP00" | AGYSUB_next == "EP00")
            
        } else {
            df = fed %>%
            filter(yr == input$year) %>%
            filter(n >= input$nmin) %>%
            filter(n_next >= input$send[1], n_next <= 2000)
        }
    
        
        sources = df %>%
            distinct(AGYSUB) %>%
            rename(label = AGYSUB)
        
        destinations = df %>%
            distinct(AGYSUB_next) %>%
            rename(label = AGYSUB_next)
        
        nodes = full_join(sources, destinations, by = "label") %>%
            rowid_to_column("id")
        
        edges = df %>%
            select(AGYSUB, AGYSUB_next, n_next) %>%
            rename(from = AGYSUB, to = AGYSUB_next, weight = n_next)
        
        nodes_d3 = mutate(nodes, id = id - 1) 
        
        nodevars = fed %>%
            group_by(AGYSUB) %>%
            slice(1) %>%
            select(AGYSUB, n, NAME)
        
        nodes_d3 = nodes_d3 %>%
            left_join(nodevars, by = c("label" = "AGYSUB")) 
        
        nodes_d3$n = nodes_d3$n %>% replace_na(500)
        
        nodes_d3[nodes_d3$label == "EXIT", "NAME"] = "EXIT" 
        
        nodes_d3$NAME = nodes_d3$NAME %>% replace_na("UNKNOWN")
        
        nodes_d3 = nodes_d3 %>%
            mutate(logn = sqrt(n))
        
        edges = left_join(edges, nodes, by = c("from" = "label")) %>%
            rename(fromid = id) %>%
            left_join(nodes, by = c("to" = "label")) %>%
            rename(toid = id) %>%
            select(fromid, toid, weight) %>%
            mutate(weight10 = sqrt(weight))
        
        edges_d3 = mutate(edges, fromid = fromid - 1, toid = toid - 1) 
        
        
        forceNetwork(Links = edges_d3, Nodes = nodes_d3, Source = "fromid", Target = "toid", 
                     NodeID = "NAME", Group = "id", Value = "weight10", 
                     opacity = 1, fontSize = 16, zoom = FALSE, Nodesize = "logn",
                     linkDistance = 100, height = 1000, arrows = TRUE)
    })

```

# ERGM

## Simplified

```{r}

########### Data Cleaning

df = ref %>%
  filter(yr != 1973) %>%
  filter(yr >= 1974) %>%
  filter(yr < 2014) %>% 
  filter(med_sal_ > 0) %>% 
  filter(n > 5000) %>%
  drop_na(med_sal_) %>%
  filter(AGYSUB != "TOTL")%>%
  filter(agy_full != "Dept. of Veterans Affairs") %>%
  filter(!grepl("NET", agy_full))

years = seq(1980,1990, by = 1)

addon = df %>%
  select(agy_full, med_sal_, agy_typ, AGYSUB, yr) %>%
  filter(yr == years[length(years)])

res = df %>%
  filter(yr %in% years) %>%
  group_by(agy_full) %>%
  summarise(ma_diff = ma_pct[which.max(yr)] - ma_pct[which.min(yr)],
            phd_diff = doc_pct[which.max(yr)] - doc_pct[which.min(yr)],
            budget_diff = (b18_roll[which.max(yr)] - b18_roll[which.min(yr)]) / b18_roll[which.min(yr)],
            dbudget_diff = (b18_d_roll[which.max(yr)] - b18_d_roll[which.min(yr)]) / b18_d_roll[which.min(yr)],
            nyear = n()) %>%
  ungroup() %>%
  filter(nyear == length(years))

fres = res %>%
  left_join(addon)


####################################################
#### NODES
######################################
#### Data Cleaning
df = fed %>%
  filter(yr == 1990) %>%
  filter(n_nextpct <= .5)


#### Make nodes
sources = df %>%
  distinct(AGYSUB) %>%
  rename(label = AGYSUB)

destinations = df %>%
  distinct(label = AGYSUB_next)

nodes = full_join(sources, destinations, by = "label")

nodes = fres %>%
  left_join(nodes, by = c("AGYSUB" = "label"))

nodes = nodes %>%
  rowid_to_column("id")


###################################3
#### Make edges
#######################


agy_set = expand.grid(unique(df$AGYSUB), unique(df$AGYSUB_next))

agy_set = agy_set %>%
  filter(as.character(Var1) != as.character(Var2)) %>%
  rename(AGYSUB = Var1, AGYSUB_next = Var2)

df2 = agy_set %>%
  left_join(df) %>%
  mutate(connected = ifelse(n_next >= 5, 1, 0))

df2$connected = replace_na(df2$connected, 0)

df2 = df2 %>%
  select(AGYSUB, AGYSUB_next, connected) %>%
  filter(connected == 1) %>%
  filter(AGYSUB %in% nodes$AGYSUB & AGYSUB_next %in% nodes$AGYSUB)

#### Make network

fednetwork = network(df2, vertex.attr = nodes, matrix.type = "edgelist")

```

```{r, results = 'asis'}

set.seed(442210)

render.d3movie(fednetwork, 
               plot.par=list(displaylabels=T, label.cex = .45), 
               label = network.vertex.names(fednetwork), 
               render.par=list(tween.frames=10,
                               show.time=TRUE,
                               show.stats= "~edges+triangles",
                               extraPlotCmds=NULL,
                               initial.coords=0),
               d3.options = list(animateOnLoad = FALSE, animationDuration = 3000,
                                 durationControl=TRUE),
               output.mode = 'inline')


```

## Simple ERGM Code

```{r, echo = TRUE}

set.seed(442210)
model1 = ergm(fednetwork ~ edges, 
             control=control.ergm(
               MCMC.samplesize=20000,
               MCMC.burnin=500000,
               MCMLE.maxit=10,
               parallel = 4))

```

## Summary Null Model

```{r}

summary(model1)

```


## Simple ERGM add covariate

```{r, echo = TRUE}

set.seed(442210)
model2 = ergm(fednetwork ~ edges + nodeicov("budget_diff"), 
             control=control.ergm(
               MCMC.samplesize=20000,
               MCMC.burnin=500000,
               MCMLE.maxit=10,
               parallel = 4))

```

## Summary Model 2

```{r}

summary(model2)

```

## Add Another Node Cov

```{r, echo = TRUE}

set.seed(442210)
model3 = ergm(fednetwork ~ edges + nodeicov("budget_diff") + nodematch("agy_typ") + absdiff("med_sal_"), 
             control=control.ergm(
               MCMC.samplesize=20000,
               MCMC.burnin=500000,
               MCMLE.maxit=10,
               parallel = 4))

```

## Summary Model3

```{r}

summary(model3)

```

## Add endog

```{r}

set.seed(442210)
model4 = ergm(fednetwork ~ edges + 
                nodeicov("budget_diff") + 
                nodematch("agy_typ") + 
                gwidegree(2), 
             control=control.ergm(
               MCMC.samplesize=20000,
               MCMC.burnin=500000,
               MCMLE.maxit=10,
               parallel = 4))

```


```{r}

summary(model4)

```
## More complex

```{r, echo = TRUE}

# set.seed(442210)
# model1 = ergm(fednetwork ~ edges + nodecov("med_sal_") +
#                 nodematch("agy_typ") +
#                 nodecov("n") +
#                 mutual + gwidegree(2),
#              control=control.ergm(
#                MCMC.samplesize=20000,
#                MCMC.burnin=500000,
#                MCMLE.maxit=10,
#                parallel = 4))

```

## Results

```{r}

# gof1 = gof(model1)
# 
# par(mfrow=c(3,2))
# plot(gof1)


```

## MCMC Diagnostics

```{r, results='hide'}

# mcmc.diagnostics(model1)

```

## TERGM

```{r}


years = seq(1974, 2012, by = 1)

nets = list()

df = fed %>%
  filter(n >= 5000) %>%
  filter(n <= 20000) %>%
  filter(n_nextpct <= .5)

sources = df %>%
  distinct(NAME) %>%
  rename(label = NAME)

destinations = df %>%
  distinct(label = NAME_next)

nodes = full_join(sources, destinations, by = "label")

nodes = nodes %>%
  rowid_to_column("id")

#######################################################

for (i in 1:length(years)) {
  
  df = fed %>%
    filter(yr == years[i]) %>%
    filter(n >= 1000) %>%
    filter(n <= 20000) %>%
    filter(n_nextpct <= .5)
  
  #### Add covariates to nodes
  
  ref2 = ref %>%
    filter(yr == years[i]) %>%
    select(n, agy_typ, med_sal_, agy_full) %>%
    mutate(agy_full = toupper(agy_full))
  
  nodestemp = nodes %>%
    left_join(ref2, by = c("label" = "agy_full")) %>%
    drop_na()
  
  #### Make edges
  
  agy_set = expand.grid(unique(df$NAME), unique(df$NAME_next))
  
  agy_set = agy_set %>%
    filter(as.character(Var1) != as.character(Var2)) %>%
    rename(NAME = Var1, NAME_next = Var2)
  
  df2 = agy_set %>%
    left_join(df) %>%
    mutate(connected = ifelse(n_next > 10, 1, 0)) 
  
  df2$connected = replace_na(df2$connected, 0)
  
  df2 = df2 %>%
    select(NAME, NAME_next, connected)
  
  df2 = df2 %>%
    left_join(nodestemp, by = c("NAME" = "label")) %>%
    rename(from = id)
  
  df2 = df2 %>%
    left_join(nodestemp, by = c("NAME_next" = "label")) %>%
    rename(to = id)
  
  df2 = df2 %>%
    filter(from != to) %>%
    filter(connected == 1)
  
  #### Make network
  
  temp = network(df2, vertex.attr = nodestemp, matrix.type = "edgelist")
  nets[[i]] = temp
  
  
}

sampdyn = networkDynamic(network.list = nets, vertex.pid = "vertex.names")



```

## Tergm2

```{r, results='asis'}

render.d3movie(sampdyn, 
               plot.par=list(displaylabels=T, label.cex = .45), 
               label = network.vertex.names(sampdyn), 
               render.par=list(tween.frames=10,
                               show.time=TRUE,
                               show.stats= "~edges+triangles",
                               extraPlotCmds=NULL,
                               initial.coords=0),
               d3.options = list(animateOnLoad = FALSE, animationDuration = 3000,
                                 durationControl=TRUE),
               output.mode = 'inline')

```



---